Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues

### 2015-02-15 android-stacks.tex
new title and bytecode explanation

\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }

@@

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analyze the source code of Java Exception classes (when the
bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.

@@

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either on Android  Aplication Framework or on third-party libraries) 
explicitly signaling a runtime exception.

Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
without listing such exception on the \emph{exception interface} (i.e., using 
throws clause on method signature). From this set only one method (defined on a library)
included an @throws tag on Javadoc - reporting that the given runtime exception
could be thrown in some contidions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other two studies~\cite{sacramento2006unchecked,kechagia2014}. 
One conducted by  Sacramento et al~\cite{sacramento2006unchecked} - which observed that the
runtime exceptions in .Net programs are most often not documented. And other conducted by 
Kechagia and Spinellis~\cite{kechagia2014} which discovered a set of methods 
on the Android API which do not documented runtime exceptions. One limitation 
of the work of Kechagia and Spinellis is that it did not filter out exceptions that,  
although runtime, should not be documented because they were implicitly signaled by the 
JVM due to resource restrictions or violations on Java semantic constraints.

When explicitly signaling a runtime exception and not documenting it, 
the developer imposes a threat to system robustness, specially
when such exceptions are thrown by a third party code (e.g. libraries, or framework utility code)
 invoked inside the application. In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

>>>>

%\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }

\title{Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues }

@@

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace needs to be classified according to its type
(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
a base exception type. Hence in this study the bytecode analysis was only used to discover the type 
of each mined exception when the jar file of such exception was available on the project or 
on the reused core Java libraries. An specific implementation (based on source code analysis)
 was needed to to discover the exception type when the bytecode was not available.

@@

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
 \emph{explicitly thrown} runtime exceptions 
should be documented as part of the \emph{exception interface} of libraries/framework reusable
methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
 signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected
the code for each method 
(defined either on Android  Aplication Framework or on third-party libraries) 
explicitly signaling a runtime exception. Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
without listing such exception on the \emph{exception interface} (i.e., using 
throws clause on method signature). From this set only one method (defined on a library)
included an @throws tag on Javadoc - reporting that the given runtime exception
could be thrown in some contidions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other two studies~\cite{sacramento2006unchecked,kechagia2014}. 
One conducted by  Sacramento et al~\cite{sacramento2006unchecked} - which observed that the
runtime exceptions in .Net programs are most often not documented. And other conducted by 
Kechagia and Spinellis~\cite{kechagia2014} which discovered a set of methods 
on the Android API which do not documented runtime exceptions. One limitation 
of the work of Kechagia and Spinellis is that it did not filter out exceptions that,  
although runtime, should not be documented because they were implicitly signaled by the 
JVM due to resource restrictions or violations on Java semantic constraints.
When explicitly signaling a runtime exception and not documenting it, 
the developer imposes a threat to system robustness, specially
when such exceptions are thrown by a third party code (e.g. libraries, or framework utility code)
 invoked inside the application. In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.
